\documentclass{beamer}
\usetheme{Madrid}
\usepackage{color}
\usecolortheme{rose}
\title{New Generation Innovation and Enterpreneuship Development Centre}
\author{Manpreet Dhiman}
\institute{G.N.D.E.C, LDH}
\date{May 3, 2016}
\begin{document}
\begin{frame}
\titlepage
\end{frame}

\begin{frame}
\frametitle{overview}
\tableofcontents
\end{frame}
\section{Start Up}
\section{IEDC}
\begin{frame}
\frametitle{Objectives of NewGen IEDC}
\begin{itemize}
\item To channelize the knowledge and the energy of youth.
\item To build a culture of innovation driven entrepreneurship.
\item To enhance the economc development.
\item Institutional mechanism for providing various services regarding startups.
\end{itemize}
\end{frame}
\begin{frame}
\frametitle{Functions of NewGen IEDC}
\begin{itemize}
\item Encouraging Students.
\item Organising Business Plans.
\item Developing innovative ideas.
\item Initiating innovative student projects.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Criteria for Selection}
\begin{itemize}
\item Institute should offer Engineering, Technology, Science courses at degree level.
\item Availability of two faculty cleared with Faculty Development Program.
\end{itemize}
\end{frame}
\begin{frame}
\frametitle{Important Forms}
\begin{itemize}
\item Main Proforma(Submission of Proposal).
\item Proforma for Proposal to be Submitted by Institutions willing to set up NewGen IEDC.
\item Guidelines for Submission of Details of Student Project.
\item The Financial Requirement.
\end{itemize}
\end{frame}

\begin{frame}
\Huge \color{blue}{\centerline {Thank You}}
\end{frame}


\end{document}